### 2019-05-23 2_metg.tex
More text.

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red lines in Figure~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s for this configuration of
Task Bench. We use 50\% because that is generally an acceptable level
of efficiency in practice, and values above 90\% can misrepresent the
performance of some systems (see
Section~\ref{subsec:peak-performance-and-efficiency}).

@@

METG has a well-defined
relationship with quantities of interest to application developers,
namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling behavior of the MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
figures, the vertical axis is shown as wall time to emphasize the
relationship to time-to-solution, but it could equivalently be shown
as task granularity (as the number of tasks per execution is
fixed). Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes dominated by overhead. The shape of the curve at the bottom is the same
for both weak and strong scaling; this is
the METG curve.

METG therefore has a direct relationship with the smallest problem
size that can be weak scaled to a given node count with a given level
of efficiency. With smaller problem sizes runtime overhead increasingly dominates useful
computation. Similarly, METG corresponds to the point at which strong
scaling can be expected to stop; as strong scaling runs start with a
larger problem size and progressively shrink it until overheads begin
to dwarf the gains due to continued scaling. In
Figures~\ref{fig:weak-scaling-mpi} and \ref{fig:strong-scaling-mpi},
the dashed, bright red line show the value of METG(50\%) at each node
count.

>>>>

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figure~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s for this configuration of
Task Bench. We use 50\% because that is generally an acceptable level
of efficiency in practice, and values above 90\% can misrepresent the
performance of some systems (see
Section~\ref{subsec:peak-performance-and-efficiency}).

@@

METG has a well-defined
relationship with quantities of interest to application developers,
namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling behavior of the MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
figures, the vertical axis is shown as wall time to emphasize the
relationship to time-to-solution, but it could equivalently be shown
as task granularity (as the number of tasks per execution is
fixed). Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes
dominated by overhead. The red, dashed line in each graph marks
METG(50\%), the point where efficiency reaches 50\%. Note that the
shapes at the bottom of the strong and weak scaling curves are identical,
and conform to the shape of the METG curve.

METG therefore has a direct relationship with the smallest problem
size that can be weak scaled to a given node count with a given level
of efficiency. Using the formula for task granularity above, each run
is 32 tasks wide and 1000 timesteps long, so task granularity is wall
time divided by 1000 (since Cori has 32 cores per node). The $2^{12}$
problem size in Figure~\ref{fig:weak-scaling-mpi} scales well
initially because the task granularity of 20~\textmu{}s is greater
than the METG(50\%) of MPI at small node counts (which is about 4.6-12
\textmu{}s from 1-64 nodes) but not at higher node counts (which rises
to 28 \textmu{}s at 128 nodes and 61 \textmu{}s at 256). Similarly,
METG corresponds to the point at which strong scaling can be expected
to stop. In Figure~\ref{fig:strong-scaling-mpi} the problem size
$2^{18}$ begins to peter out at 64 nodes, the point at which the
scaling curve intersects METG(50\%).
